---
title: "Genera_Level_Population_Analysis"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Import Packages
```{r}
library("vegan")
library("calibrate")
library("dplyr")
library("tidyverse")
```
Hypotheses: 
1. Populations differ significantly by test (APC, GN)
2. Populations differ significantly by sampling (S1, S2, S3) 
3. Populations differ significantly by day of isolate collection (D0, D3, D10)

Aim: 
1. Check whether it is appropriate to use a Chi-squared test, using the coarsest division of the data (by test) (i.e., whether more than 20% of the cells have an expected value greater than 5)
2. If chi-squared tests are not appropriate, check whether it is okay to use a Fisher's exact test with the simulate.p.value function for tables larger than 2x2
3. Use Kruskal-Wallis tests, with species richness 

Guide to dfs:
df_1 <- raw data of the isolates with Sequence_Name (original_ID & blast identity concatenated) and the ST assignment
df_2 <- Split the Sequence_Name in df_1 into the Original ID, Genus and Species. Kept the Sequence_Name vector
df_3 <- Split the Original_ID to get the Location, Sampling, Day, BiologicalRep, Test, TechnicalRep and IsolateNumber vectors. Kept the Original_ID vector. 
df_4 <- Isolates from APC and GN tests. Created a new column: High_Frequency_Genus. Genera with less than 3 isolates among APC and GN isolates were labeled as "Rare"
df_5 <- Isolates from APC, subset of df_4
df_6 <- Isolates from GN,  subset of df_4
df_7 <- Isolates from BHI, subset of df_2. 

test_ct <- Contingency table of High_Frequency_Genus and Test using data from df_4 
sampling_ct <- Contingency table of High_Frequency_Genus and Sampling using data from df_4 
day_ct <- Contingency table of High_Frequency_Genus and Day using data from df_4 
apc_sampling_ct <- Contingency table of High_Frequency_Genus and Sampling using data from df_5
apc_day_ct <- Contingency table of High_Frequency_Genus and Day using data from df_5
gn_sampling_ct <- Contingency table of High_Frequency_Genus and Sampling using data from df_6 
gn_day_ct <- Contingency table of High_Frequency_Genus and Day using data from df_6 
bhi_sampling_ct <- Contingency table of High_Frequency_Genus and Sampling using data from df_7
bhi_day_ct <- Contingency table of High_Frequency_Genus and Day using data from df_7

Analysis overview 
Analysis for BHI and APC/GN data conducted separately due to difference in methodology for collecting and sequencing isolates. 
APC/GN data:
Assign any genera with less than three isolates as "Rare". 
Test whether there is an association of genera by test. If so, analyze APC and GN data separately. If not, test all petrifilm data (APC and GN) together
Conduct Fisher's exact tests to look for association of genera to: i) sampling, ii) day 

BHI data:
Assign any genera with less than three isolates as "Rare". 
Conduct Fisher's exact tests to look for association of genera to: i) sampling, ii) day 

Chunk Description: Parsing dataframes to aid in analysis. This chunk uses functions from the tidyr package (separate) and the dplyr package (mutate)
Chunk conclusion: NA

Key:
sampling = lot, therefore S1=L1, S2=L2, S3=L3
BJ refers to Beijing eCommerce distribution pathway, H/HD refers to local grocery distribution pathway
```{r}
df_1 <- read.csv("df_isolateInfo_051321.csv", header = TRUE)

df_2 <- df_1 %>%
  separate(isolateID, c("Location", "Sampling", "Day", "BiologicalRep", "Test", "TechnicalRep", "IsolateNumber"), sep = "-", extra = "merge", remove = FALSE)

##Edit elements in dataframe to retain consistency.
#df_3$Location, changed HD to H 

df_2 <- df_2 %>%
  mutate(Location = replace(Location, Location == "HD", "H"))

#df_3$Day, changed DI to D3
df_2 = df_2 %>%
  mutate(Day = replace(Day, Day == "DI", "D3"))

df_3 <- df_2 %>%
  filter(Test != "BHI")

isolates_summary_df_3 <- data.frame(table(df_3$genus_C))

rare_isolates_df_3 <- isolates_summary_df_3 %>%
  filter(Freq < 3) 

df_4 <- df_3 %>%
  mutate(High_Frequency_Genus =  replace(genus_C, genus_C %in% rare_isolates_df_3$Var1, "Rare"))

df_5 <- df_4 %>%
  filter(Test == "APC")

df_6 <- df_4 %>%
  filter(Test == "GN")

df_7 <- df_2 %>%
  filter(Test == "BHI")

isolate_summary_df_7 <- data.frame(table(df_7$genus_C))

rare_isolates_df_7 <- isolate_summary_df_7 %>%
  filter(Freq < 3) 

df_7 <- df_7 %>%
  mutate(High_Frequency_Genus =  replace(genus_C, genus_C %in% rare_isolates_df_7$Var1, "Rare"))
```

Chunk Description: Conducting a chi-squared test of genera by test (APC and GN) (coarsest sub-division of the data). If expected values for more than 20% of the cells is greater than 5, we can use this test. This is only for isolates selected from APC and GN petrifilms 
Conclusion:The expected values of >20% cells is less than 5 for sampling_ct (APC and GN data) and day_ct (APC and GN data). Hence, we cannot use the chi-squared test (this is when the APC and GN data were combined. When we analyze the APC and GN data separately, a greater proportion of the cells have expected values <5)
```{r}
test_ct <- table(df_4$High_Frequency_Genus, df_4$Test)
test_chisq <- chisq.test(test_ct)
test_chisq$expected

sampling_ct <- table(df_4$High_Frequency_Genus, df_4$Sampling)
sampling_chisq <- chisq.test(sampling_ct)
sampling_chisq$expected

day_ct <- table(df_4$High_Frequency_Genus, df_4$Day)
day_chisq <- chisq.test(day_ct)
day_chisq$expected
```

```{r}
sink("Fishers_Exact_Test_Output.txt")

print("Hypotheses:", quote = FALSE) 
print("1. Populations differ significantly by test (APC, GN)", quote = FALSE)
print("2. Populations differ significantly by sampling (S1, S2, S3)", quote = FALSE)
print("3. Populations differ significantly by day of isolate collection (D0, D3, D10)", quote = FALSE)

print("Aim: Use Fisher's exact test with the simulate.p.value function for tables larger than 2x2, to test for significant differences in isolate identity between: 1) test, 2) sampling, 3)day of testing/isolate collection", quote = FALSE)

print("1) Test of significant difference between tests, for the contingency table below (Hypothesis 1):", quote = FALSE)

print("Test_Contingency_Table", quote = FALSE)
print(test_ct)

print("Result of Fisher's Exact Test for Hypothesis 1:", quote= FALSE)
fisher.test(test_ct, simulate.p.value = TRUE)

print("Sampling_Contingency_Table", quote = FALSE)
print(sampling_ct)

print("Result of Fisher's Exact Test for Hypothesis 2:", quote= FALSE)
fisher.test(sampling_ct, simulate.p.value = TRUE)

print("Day_Contingency_Table", quote = FALSE)
print(day_ct)

print("Result of Fisher's Exact Test for Hypothesis 3:", quote= FALSE)
fisher.test(day_ct, simulate.p.value = TRUE)
sink()
```

```{r}
apc_sampling_ct <- table(df_5$High_Frequency_Genus, df_5$Sampling)

apc_day_ct <- table(df_5$High_Frequency_Genus, df_5$Day)

gn_sampling_ct <- table(df_6$High_Frequency_Genus, df_6$Sampling)

gn_day_ct <- table(df_6$High_Frequency_Genus, df_6$Day)

```

Fisher's exact test for APC data: association of genera for: i) lot, ii) day 
```{r}
sink("Fishers_Exact_Test_Output_APC_Only.txt")

print("Test of significant difference between samplings, for the contingency table below (Hypothesis 1; APC data only):", quote = FALSE)

print("Sampling_ContingencyTable_APC", quote = FALSE)
print(apc_sampling_ct)

print("Result of Fisher's Exact Test for Sampling_ContingencyTable_APC:", quote= FALSE)
fisher.test(apc_sampling_ct, simulate.p.value = TRUE)

print("Day_ContingencyTable_APC", quote = FALSE)
print(apc_day_ct)

print("Result of Fisher's Exact Test for Day_ContingencyTable_APC:", quote= FALSE)
fisher.test(apc_day_ct, simulate.p.value = TRUE)

sink()
```


Fisher's exact test for GN data: association of genera for: i) lot, ii) day 
```{r}
sink("Fishers_Exact_Test_Output_GN_Only.txt")

print("Test of significant difference between samplings, for the contingency table below (Hypothesis 1; APC data only):", quote = FALSE)

print("Sampling_ContingencyTable_GN", quote = FALSE)
print(gn_sampling_ct)

print("Result of Fisher's Exact Test for Sampling_ContingencyTable_GN:", quote= FALSE)
fisher.test(gn_sampling_ct, simulate.p.value = TRUE)

print("Day_ContingencyTable_GN", quote = FALSE)
print(gn_day_ct)

print("Result of Fisher's Exact Test for Day_ContingencyTable_GN:", quote= FALSE)
fisher.test(gn_day_ct, simulate.p.value = TRUE)

sink()
```

```{r}
bhi_sampling_ct <- table(df_7$High_Frequency_Genus, df_7$Sampling)

bhi_day_ct <- table(df_7$High_Frequency_Genus, df_7$Day)
```

Fisher's exact test for BHI data: association of genera for: i) lot, ii) day 
```{r}
sink("Fishers_Exact_Test_Output_BHI_Only.txt")

print("Test of significant difference between samplings, for the contingency table below (Hypothesis 1; BHI data only):", quote = FALSE)

print("Sampling_ContingencyTable_BHI", quote = FALSE)
print(bhi_sampling_ct)

print("Result of Fisher's Exact Test for Sampling_ContingencyTable_BHI:", quote= FALSE)
fisher.test(bhi_sampling_ct, simulate.p.value = TRUE)

print("Day_ContingencyTable_BHI", quote = FALSE)
print(bhi_day_ct)

print("Result of Fisher's Exact Test for Day_ContingencyTable_BHI:", quote= FALSE)
fisher.test(bhi_day_ct, simulate.p.value = TRUE)

sink()
```




