---
title: "DataAnalysis"
author: "Sarah I. Murphy"
output: 
  html_document:
      toc: true
      number_sections: true
      toc_float: true
      code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file contains analysis for logistics, environmental, and microbiological (APC & GN) data. Abbreviations are sometimes used where "E" represents Beijing eCommerce, "L" represents Local grocery store, "GH" represents greenhouse, "P" represents processing facility.

```{r, warning = FALSE, message=FALSE}
# load packages
library(readr)
library(tidyr)
library(dplyr)
library(stats)
library(lme4)
library(ggplot2)
library(ggpubr)
```

# 1. Logistics 

## Timing of spinach supply chain logistics for lots 1, 2, and 3
```{r}
# load data
df_timing <- read_csv("Data/LogisticsTimingData.csv")
# prep table
df_timing %>%
  group_by(lot)%>%
  spread(lot,date_time)-> TableS1

# clean-up environment
rm(df_timing) 
```

## Summary stats of supply chain activity duration
```{r,warning = FALSE, message=FALSE}
# load data
df_activity <-read_csv("Data/LogisticsActivityDurationData.csv") 

# summarize median, minimum, and maximum duration for each activity
df_activity %>%
  group_by(Activity) %>%
  summarize(                
    median = median(Duration_minutes, na.rm = TRUE) / 60, #convert from minutes to hours by dividing by 60
    min = min(Duration_minutes, na.rm = TRUE) / 60,
    max = max(Duration_minutes, na.rm = TRUE) / 60,
    n = sum(!is.na(Duration_minutes))) -> summary_duration

# how much overlap was there between cold storage for E and L samples?
df_activity %>%
  filter(Activity == "PackagedProductForE_ColdStorage" |
      Activity == "PackagedProductForL_ColdStorage") %>%
  group_by(Lot) %>%
  spread(., Activity, Duration_minutes) %>%
  group_by(Lot) %>%
  summarize(durationDiff_h = (
      PackagedProductForE_ColdStorage - PackagedProductForL_ColdStorage)/60
  ) -> summary_duration_csOverlap

#rename df_activity as Table1a
Table1_duration <- df_activity

#clean-up environment
rm(df_activity)
```
# 2. Environmental conditions data

## a. Environmental conditions across spinach supply chain activities (dataloggers)

## Summary stats spinach supply chain environmental conditions
```{r,warning = FALSE, message=FALSE}
# prep data
dfl <- read_csv("Data/SupplyChainLoggerData.csv")

# Temperature summary by activity & lot
dfl %>%
  filter(is.na(activity)!=TRUE)%>%
  filter(lot!="L1"|activity!="3_RawProduct_TransitGHtoP")%>%#missing transit start time 
  group_by(lot, activity) %>%
  summarize(mean = mean(Temp_C),
            sd = sd(Temp_C),
            min = min(Temp_C),
            quantile_25 = quantile(Temp_C,0.25),
            median = median(Temp_C),
            quantile_75 = quantile(Temp_C,0.75),
            max = max(Temp_C),
            n_datapoints=n()) ->Table1_Temp

# Relative humidity summary by activity & lot
dfl %>%
  filter(is.na(RH_Percent)!=TRUE) %>% #some don't have RH_Percent data
  filter(is.na(activity)!=TRUE)%>%
  filter(lot!="L1"|activity!="3_RawProduct_TransitGHtoP")%>%#missing transit start time
  group_by(lot, activity) %>%
  summarize(mean = mean(RH_Percent),
            sd=sd(RH_Percent),
            min = min(RH_Percent),
            quantile_25 = quantile(RH_Percent,0.25),
            median = median(RH_Percent),
            quantile_75 = quantile(RH_Percent,0.75),
            max = max(RH_Percent),
            n_datapoints=n()) -> Table1_RH
```

## Plot spinach supply chain environmental conditions data over time
```{r}
#plot temperature
t_L1 <- dfl %>%
  filter(lot=="L1") %>%
  filter(activity!="3_RawProduct_TransitGHtoP")%>%#missing transit start time
  ggplot(., aes(x=China_DateTime,y=Temp_C,col=activity))+
  geom_point(shape=1)+
  scale_color_manual(values=c("#00BA38", "#00C19F", "#00B9E3",
                              "#619CFF", "#DB72FB","#FF61C3"))+
  scale_y_continuous(breaks=seq(-2,15,1),limits=c(-2,15))+
  scale_x_datetime(breaks = seq(as.POSIXct("0020-01-06 12:00:00"),
                                as.POSIXct("0020-01-10 12:00:00"), "24 hour"),
                   date_labels = "%b %d - %H:%M",
                   limits = c(as.POSIXct("0020-01-06 09:00:00"),
                              as.POSIXct("0020-01-10 12:00:00")))+
  xlab("")+ ylab("Temperature (°C)")+
  ggtitle("L1")+
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 25, vjust = 1.0, hjust = 1.0))

t_L2 <- dfl %>%
  filter(lot=="L2") %>%
  ggplot(., aes(x=China_DateTime,y=Temp_C,col=activity))+
  geom_point(shape=1)+
  scale_y_continuous(breaks=seq(-2,15,1),limits=c(-2,15))+
  scale_x_datetime(breaks = seq(as.POSIXct("0020-01-07 12:00:00"),
                                as.POSIXct("0020-01-11 12:00:00"), "24 hour"),
                   date_labels = "%b %d - %H:%M",
                   limits = c(as.POSIXct("0020-01-07 09:00:00"),
                              as.POSIXct("0020-01-11 12:00:00")))+
  xlab("")+ ylab("Temperature (°C)")+
  ggtitle("L2")+
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 25, vjust = 1.0, hjust = 1.0))

t_L3 <- dfl %>%
  filter(lot=="L3") %>%
  ggplot(., aes(x=China_DateTime,y=Temp_C,col=activity))+
  geom_point(shape=1)+
  scale_y_continuous(breaks=seq(-2,15,1),limits=c(-2,15))+
  scale_x_datetime(breaks = seq(as.POSIXct("0020-01-08 12:00:00"),
                                as.POSIXct("0020-01-12 12:00:00"), "24 hour"),
                   date_labels = "%b %d - %H:%M",
                   limits = c(as.POSIXct("0020-01-08 09:00:00"),
                              as.POSIXct("0020-01-12 12:00:00")))+
  xlab("")+ 
  ylab("Temperature (°C)")+
  ggtitle("L3")+
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 25, vjust = 1.0, hjust = 1.0))

#plot rh
rh_L1 <- dfl %>%
  filter(lot=="L1") %>%
  filter(activity!="3_RawProduct_TransitGHtoP")%>%#missing transit start time 
  filter(RH_Percent!="NA")%>%
  ggplot(., aes(x=China_DateTime,y=as.numeric(RH_Percent),col=activity))+
  geom_point(shape=1)+
  scale_color_manual(values=c( "#00BA38", "#00C19F", "#00B9E3",
                              "#619CFF", "#DB72FB","#FF61C3"))+
  scale_y_continuous(limits=c(0,100))+
  scale_x_datetime(breaks = seq(as.POSIXct("0020-01-06 12:00:00"),
                                as.POSIXct("0020-01-10 12:00:00"), "24 hour"),
                   date_labels = "%b %d - %H:%M",
                   limits = c(as.POSIXct("0020-01-06 09:00:00"),
                              as.POSIXct("0020-01-10 12:00:00")))+
  xlab("")+ 
  ylab("Relative Humidity (%)")+
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 25, vjust = 1.0, hjust = 1.0))

rh_L2 <- dfl %>%
  filter(lot=="L2") %>%
  filter(RH_Percent!="NA")%>%
  ggplot(., aes(x=China_DateTime,y=as.numeric(RH_Percent),col=activity))+
  geom_point(shape=1)+
  scale_y_continuous(limits=c(0,100))+
  scale_x_datetime(breaks = seq(as.POSIXct("0020-01-07 12:00:00"),
                                as.POSIXct("0020-01-11 12:00:00"), "24 hour"),
                   date_labels = "%b %d - %H:%M",
                   limits = c(as.POSIXct("0020-01-07 09:00:00"),
                              as.POSIXct("0020-01-11 12:00:00")))+
  xlab("")+ 
  ylab("Relative Humidity (%)")+
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 25, vjust = 1.0, hjust = 1.0))

rh_L3 <- dfl %>%
  filter(lot=="L3") %>%
  filter(RH_Percent!="NA")%>%
  ggplot(., aes(x=China_DateTime,y=as.numeric(RH_Percent),col=activity))+
  geom_point(shape=1)+
  scale_y_continuous(limits=c(0,100))+
  scale_x_datetime(breaks = seq(as.POSIXct("0020-01-08 12:00:00"),
                                as.POSIXct("0020-01-12 12:00:00"), "24 hour"),
                   date_labels = "%b %d - %H:%M",
                   limits = c(as.POSIXct("0020-01-08 09:00:00"),
                              as.POSIXct("0020-01-12 12:00:00")))+
  xlab("")+ 
  ylab("Relative Humidity (%)")+
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 25, vjust = 1.0, hjust = 1.0))

# arrange plots
ggarrange(t_L1, t_L2, t_L3, rh_L1, rh_L2,rh_L3,ncol=3,nrow=2, align="v") -> plot_enviro_supplychain

# clean-up environment
rm(t_L1, t_L2, t_L3, rh_L1, rh_L2, rh_L3, dfl)
```

## b. Retail display environmental conditions (datalogger)
```{r,warning = FALSE, message=FALSE}
dfr <- read_csv("Data/RetailDisplayLoggerData.csv")

dfr$China_DateTime = as.POSIXct(dfr$China_DateTime, format="%m/%d/%Y %H:%M")

dfr %>% separate(., China_DateTime, c("Date", "Time"), sep = " ", remove=FALSE) -> dfr
dfr %>% # restrict to when the datalogger was actually in the display case
  filter(Date!="0020-01-09"|Time>="15:10:00")->dfr
```
## Summary stats retail display environmental conditions 
```{r}
# summary stats
dfr %>% summarize(name="Temperature",
            mean=mean(`Chan 1 - Deg C`),
            sd=sd(`Chan 1 - Deg C`),
            median=median(`Chan 1 - Deg C`),
            quantile_25=quantile(`Chan 1 - Deg C`,0.25),
            quantile_75=quantile(`Chan 1 - Deg C`,0.75),
            min=min(`Chan 1 - Deg C`),
            max=max(`Chan 1 - Deg C`),
            n=n())->dfr_t

dfr %>% summarize(name="Relative Humidity",
            mean=mean(`Chan 2 - %RH`),
            sd=sd(`Chan 2 - %RH`),
            median=median(`Chan 2 - %RH`),
            quantile_25=quantile(`Chan 2 - %RH`,0.25),
            quantile_75=quantile(`Chan 2 - %RH`,0.75),
            min=min(`Chan 2 - %RH`),
            max=max(`Chan 2 - %RH`),
            n=n())->dfr_rh

rbind(dfr_t,dfr_rh)-> summary_enviro_retaildisplay

# clean-up environment
rm(dfr_t, dfr_rh)
```

## Plot retail display environmental conditions data over time 
```{r}
# plot
FigureS2_Temp <- ggplot(dfr, aes(x=China_DateTime,y=`Chan 1 - Deg C`))+
  geom_point(alpha = 0.3)+
    scale_x_datetime(breaks = seq(as.POSIXct("0020-01-09 12:00:00"),
                                  as.POSIXct("0020-01-17 12:00:00"), "24 hour"),
                     date_labels = "%b %d - %H:%M",
                     limits = c(as.POSIXct("0020-01-09 12:00:00"),
                                as.POSIXct("0020-01-17 12:00:00")))+
  xlab("")+
  ylab(expression(paste("Temperature (",degree,"C )")))+
  theme(axis.text.x = element_text(size=12),axis.text.y = element_text(size=12),axis.title.y = element_text(size=12),title = element_text(size=14))

FigureS2_RH <- ggplot(dfr, aes(x=China_DateTime,y=`Chan 2 - %RH`))+
  geom_point(alpha = 0.3)+
    scale_x_datetime(breaks = seq(as.POSIXct("0020-01-09 12:00:00"),
                                  as.POSIXct("0020-01-17 12:00:00"), "24 hour"),
                     date_labels = "%b %d - %H:%M",
                     limits = c(as.POSIXct("0020-01-09 12:00:00"),
                                as.POSIXct("0020-01-17 12:00:00")))+
  xlab("")+
  ylab("Relative Humidity (%)")+
  theme(axis.text.x = element_text(size=12),axis.text.y = element_text(size=12),axis.title.y = element_text(size=12),title = element_text(size=14))

ggarrange(FigureS2_Temp,FigureS2_RH,ncol=1,nrow=2) ->FigureS2

# clean-up environment
rm(dfr,FigureS2_Temp,FigureS2_RH)
```

## c. Greenhouse environmental conditions (USB micro-station)
```{r}
df_enviro <- read_csv("Data/GreenhouseEnvironmentStationData.csv")
df_enviro %>% separate(., China_DateTime, c("Date", "Time"), sep = " ", remove=FALSE) -> df_enviro
#remove datapoint from start-up station
df_enviro %>%
  filter(Date!="0020-01-05"|Time!="10:51:00") -> df_enviro
# prep study period subset
df_enviro %>%
  filter(Date=="0020-01-05"|Date=="0020-01-06"|Date=="0020-01-07"|Date=="0020-01-08"&Time<="15:12:00")->df_enviro2
```

## Summary stats greenhouse environmental conditions overall
```{r}
df_enviro2 %>% summarize(name="LeafWetness_Percent",
            mean=mean(LeafWetness_Percent),
            sd=sd(LeafWetness_Percent),
            median=median(LeafWetness_Percent),
            quantile_25=quantile(LeafWetness_Percent,0.25),
            quantile_75=quantile(LeafWetness_Percent,0.75),
            min=min(LeafWetness_Percent),
            max=max(LeafWetness_Percent),
            n=n())->df_lw

df_enviro2 %>% summarize(name="SolarRadiation_WattsPerSquareMeter",
            mean=mean(SolarRadiation_WattsPerSquareMeter),
            sd=sd(SolarRadiation_WattsPerSquareMeter),
            median=median(SolarRadiation_WattsPerSquareMeter),
            quantile_25=quantile(SolarRadiation_WattsPerSquareMeter,0.25),
            quantile_75=quantile(SolarRadiation_WattsPerSquareMeter,0.75),
            min=min(SolarRadiation_WattsPerSquareMeter),
            max=max(SolarRadiation_WattsPerSquareMeter),
            n=n())->df_sr

df_enviro2 %>% summarize(name="Temp_C",
            mean=mean(Temp_C),
            sd=sd(Temp_C),
            median=median(Temp_C),
            quantile_25=quantile(Temp_C,0.25),
            quantile_75=quantile(Temp_C,0.75),
            min=min(Temp_C),
            max=max(Temp_C),
            n=n())->df_tc

df_enviro2 %>% summarize(name="RH_Percent",
            mean=mean(RH_Percent),
            sd=sd(RH_Percent),
            median=median(RH_Percent),
            quantile_25=quantile(RH_Percent,0.25),
            quantile_75=quantile(RH_Percent,0.75),
            min=min(RH_Percent),
            max=max(RH_Percent),
            n=n())->df_rh

df_enviro2 %>% summarize(name="DewPt_C",
            mean=mean(DewPt_C),
            sd=sd(DewPt_C),
            median=median(DewPt_C),
            quantile_25=quantile(DewPt_C,0.25),
            quantile_75=quantile(DewPt_C,0.75),
            min=min(DewPt_C),
            max=max(DewPt_C),
            n=n())->df_dp

rbind(df_lw,df_sr,df_tc,df_rh,df_dp)-> summary_enviro_greenhouse

# clean-up environment
rm(df_lw,df_sr,df_tc,df_rh,df_dp)
```

## Summary stats greenhouse environmental conditions by lot
```{r}
#subset 24 h prior to L1 (only have data starting at 11:06 due to set-up of station so missing approx 1 h)
df_enviro2 %>%
  filter(Date=="0020-01-05"|Date=="0020-01-06"&Time<="10:13:00")->df_enviro2_L1
df_enviro2_L1$lot <- "L1"

#subset 24 h prior to L2
df_enviro2 %>%
  filter(Date=="0020-01-06"&Time>"09:08:00"|Date=="0020-01-07"&Time<="09:08:00")->df_enviro2_L2
df_enviro2_L2$lot <- "L2"

#subset 24 h prior to L3
df_enviro2 %>%
  filter(Date=="0020-01-07"&Time>"09:25:00"|Date=="0020-01-08"&Time<="09:25:00")->df_enviro2_L3
df_enviro2_L3$lot <- "L3"

df_enviro2byLot<- rbind(df_enviro2_L1,df_enviro2_L2,df_enviro2_L3)
```

```{r}
df_enviro2byLot %>% group_by(lot) %>%
    summarize(name="LeafWetness_Percent",
            mean=mean(LeafWetness_Percent),
            sd=sd(LeafWetness_Percent),
            median=median(LeafWetness_Percent),
            quantile_25=quantile(LeafWetness_Percent,0.25),
            quantile_75=quantile(LeafWetness_Percent,0.75),
            min=min(LeafWetness_Percent),
            max=max(LeafWetness_Percent),
            n=n())->df2_lw

df_enviro2byLot %>% group_by(lot) %>%
  summarize(name="SolarRadiation_WattsPerSquareMeter",
            mean=mean(SolarRadiation_WattsPerSquareMeter),
            sd=sd(SolarRadiation_WattsPerSquareMeter),
            median=median(SolarRadiation_WattsPerSquareMeter),
            quantile_25=quantile(SolarRadiation_WattsPerSquareMeter,0.25),
            quantile_75=quantile(SolarRadiation_WattsPerSquareMeter,0.75),
            min=min(SolarRadiation_WattsPerSquareMeter),
            max=max(SolarRadiation_WattsPerSquareMeter),
            n=n())->df2_sr

df_enviro2byLot %>% group_by(lot) %>%
  summarize(name="Temp_C",
            mean=mean(Temp_C),
            sd=sd(Temp_C),
            median=median(Temp_C),
            quantile_25=quantile(Temp_C,0.25),
            quantile_75=quantile(Temp_C,0.75),
            min=min(Temp_C),
            max=max(Temp_C),
            n=n())->df2_tc

df_enviro2byLot %>% group_by(lot) %>%
  summarize(name="RH_Percent",
            mean=mean(RH_Percent),
            sd=sd(RH_Percent),
            median=median(RH_Percent),
            quantile_25=quantile(RH_Percent,0.25),
            quantile_75=quantile(RH_Percent,0.75),
            min=min(RH_Percent),
            max=max(RH_Percent),
            n=n())->df2_rh

df_enviro2byLot %>% group_by(lot) %>%
  summarize(name="DewPt_C",
            mean=mean(DewPt_C),
            sd=sd(DewPt_C),
            median=median(DewPt_C),
            quantile_25=quantile(DewPt_C,0.25),
            quantile_75=quantile(DewPt_C,0.75),
            min=min(DewPt_C),
            max=max(DewPt_C),
            n=n())->df2_dp

rbind(df2_lw,df2_sr,df2_tc,df2_rh,df2_dp)-> Table2

# clean-up environment
rm(df2_lw,df2_sr,df2_tc,df2_rh,df2_dp,df_enviro2_L1,df_enviro2_L2,df_enviro2_L3,df_enviro2byLot)
```

## Plot greenhouse environmental conditions data over time
```{r}
## plot
lw <- ggplot(df_enviro2, aes(x=China_DateTime,y=LeafWetness_Percent))+
  scale_x_datetime(breaks = seq(as.POSIXct("0020-01-05 12:00:00"),
                                as.POSIXct("0020-01-8 12:00:00"), "24 hour"),
                   date_labels = "%b %d - %H:%M",
                   limits = c(as.POSIXct("0020-01-05 11:00:00"),
                              as.POSIXct("0020-01-8 16:00:00")))+
  xlab("")+
  ylab("Leaf Wetness (%)")+
  ggtitle("Leaf Wetness ")+
   annotate("rect",xmin=as.POSIXct("0020-01-06 10:13:00"),
            xmax=as.POSIXct("0020-01-06 10:53:00"),
            ymin=-Inf,ymax=Inf, alpha=0.3, fill="#F8766D")+
   annotate("rect",xmin=as.POSIXct("0020-01-07 09:08:00"),
            xmax=as.POSIXct("0020-01-07 09:25:00"),
            ymin=-Inf,ymax=Inf, alpha=0.3, fill="#F8766D")+
   annotate("rect",xmin=as.POSIXct("0020-01-08 09:25:00"),
            xmax=as.POSIXct("0020-01-08 09:47:00"),
            ymin=-Inf,ymax=Inf, alpha=0.3, fill="#F8766D")+
   annotate("rect",xmin=as.POSIXct("0020-01-07 09:25:00"),
            xmax=as.POSIXct("0020-01-07 14:19"),
            ymin=-Inf,ymax=Inf, alpha=0.3, fill="#00BFC4")+
   annotate("rect",xmin=as.POSIXct("0020-01-08 09:47:00"),
            xmax=as.POSIXct("0020-01-08 15:11:00"),
            ymin=-Inf,ymax=Inf, alpha=0.3, fill="#00BFC4")+
theme(axis.text.x = element_text(size=12),axis.text.y = element_text(size=12),axis.title.y = element_text(size=12),title = element_text(size=14))+
  geom_point()

sr <- ggplot(df_enviro2, aes(x=China_DateTime,y=SolarRadiation_WattsPerSquareMeter))+
  scale_x_datetime(breaks = seq(as.POSIXct("0020-01-05 12:00:00"),
                                as.POSIXct("0020-01-8 12:00:00"), "24 hour"),
                   date_labels = "%b %d - %H:%M",
                   limits = c(as.POSIXct("0020-01-05 11:00:00"),
                              as.POSIXct("0020-01-8 16:00:00")))+
  xlab("")+
  ylab(expression(paste("Solar Radiation (watts/",m^2,")")))+
  ggtitle("Solar Radiation")+
   annotate("rect",xmin=as.POSIXct("0020-01-06 10:13:00"),
            xmax=as.POSIXct("0020-01-06 10:53:00"),
            ymin=-Inf,ymax=Inf, alpha=0.3, fill="#F8766D")+
   annotate("rect",xmin=as.POSIXct("0020-01-07 09:08:00"),
            xmax=as.POSIXct("0020-01-07 09:25:00"),
            ymin=-Inf,ymax=Inf, alpha=0.3, fill="#F8766D")+
   annotate("rect",xmin=as.POSIXct("0020-01-08 09:25:00"),
            xmax=as.POSIXct("0020-01-08 09:47:00"),
            ymin=-Inf,ymax=Inf, alpha=0.3, fill="#F8766D")+
   annotate("rect",xmin=as.POSIXct("0020-01-07 09:25:00"),
            xmax=as.POSIXct("0020-01-07 14:19"),
            ymin=-Inf,ymax=Inf, alpha=0.3, fill="#00BFC4")+
   annotate("rect",xmin=as.POSIXct("0020-01-08 09:47:00"),
            xmax=as.POSIXct("0020-01-08 15:11:00"),
            ymin=-Inf,ymax=Inf, alpha=0.3, fill="#00BFC4")+
theme(axis.text.x = element_text(size=12),axis.text.y = element_text(size=12),axis.title.y = element_text(size=12),title = element_text(size=14))+
  geom_point()

tc <- ggplot(df_enviro2, aes(x=China_DateTime,y=Temp_C))+
  scale_x_datetime(breaks = seq(as.POSIXct("0020-01-05 12:00:00"),
                                as.POSIXct("0020-01-8 12:00:00"), "24 hour"),
                   date_labels = "%b %d - %H:%M",
                   limits = c(as.POSIXct("0020-01-05 11:00:00"),
                              as.POSIXct("0020-01-8 16:00:00")))+
  xlab("")+
  ylab(expression(paste("Temperature (",degree,"C )")))+
  ggtitle("Temperature")+
   annotate("rect",xmin=as.POSIXct("0020-01-06 10:13:00"),
            xmax=as.POSIXct("0020-01-06 10:53:00"),
            ymin=-Inf,ymax=Inf, alpha=0.3, fill="#F8766D")+
   annotate("rect",xmin=as.POSIXct("0020-01-07 09:08:00"),
            xmax=as.POSIXct("0020-01-07 09:25:00"),
            ymin=-Inf,ymax=Inf, alpha=0.3, fill="#F8766D")+
   annotate("rect",xmin=as.POSIXct("0020-01-08 09:25:00"),
            xmax=as.POSIXct("0020-01-08 09:47:00"),
            ymin=-Inf,ymax=Inf, alpha=0.3, fill="#F8766D")+
   annotate("rect",xmin=as.POSIXct("0020-01-07 09:25:00"),
            xmax=as.POSIXct("0020-01-07 14:19"),
            ymin=-Inf,ymax=Inf, alpha=0.3, fill="#00BFC4")+
   annotate("rect",xmin=as.POSIXct("0020-01-08 09:47:00"),
            xmax=as.POSIXct("0020-01-08 15:11:00"),
            ymin=-Inf,ymax=Inf, alpha=0.3, fill="#00BFC4")+
theme(axis.text.x = element_text(size=12),axis.text.y = element_text(size=12),axis.title.y = element_text(size=12),title = element_text(size=14))+
  geom_point()

rh <- ggplot(df_enviro2, aes(x=China_DateTime,y=RH_Percent))+
  scale_x_datetime(breaks = seq(as.POSIXct("0020-01-05 12:00:00"),
                                as.POSIXct("0020-01-8 12:00:00"), "24 hour"),
                   date_labels = "%b %d - %H:%M",
                   limits = c(as.POSIXct("0020-01-05 11:00:00"),
                              as.POSIXct("0020-01-8 16:00:00")))+
  xlab("")+
  ylab("Relative Humidity (%)")+
  ggtitle("Relative Humidity")+
   annotate("rect",xmin=as.POSIXct("0020-01-06 10:13:00"),
            xmax=as.POSIXct("0020-01-06 10:53:00"),
            ymin=-Inf,ymax=Inf, alpha=0.3, fill="#F8766D")+
   annotate("rect",xmin=as.POSIXct("0020-01-07 09:08:00"),
            xmax=as.POSIXct("0020-01-07 09:25:00"),
            ymin=-Inf,ymax=Inf, alpha=0.3, fill="#F8766D")+
   annotate("rect",xmin=as.POSIXct("0020-01-08 09:25:00"),
            xmax=as.POSIXct("0020-01-08 09:47:00"),
            ymin=-Inf,ymax=Inf, alpha=0.3, fill="#F8766D")+
   annotate("rect",xmin=as.POSIXct("0020-01-07 09:25:00"),
            xmax=as.POSIXct("0020-01-07 14:19"),
            ymin=-Inf,ymax=Inf, alpha=0.3, fill="#00BFC4")+
   annotate("rect",xmin=as.POSIXct("0020-01-08 09:47:00"),
            xmax=as.POSIXct("0020-01-08 15:11:00"),
            ymin=-Inf,ymax=Inf, alpha=0.3, fill="#00BFC4")+
theme(axis.text.x = element_text(size=12),axis.text.y = element_text(size=12),axis.title.y = element_text(size=12),title = element_text(size=14))+
  geom_point()

dp <- ggplot(df_enviro2, aes(x=China_DateTime,y=DewPt_C))+
  scale_x_datetime(breaks = seq(as.POSIXct("0020-01-05 12:00:00"),
                                as.POSIXct("0020-01-8 12:00:00"), "24 hour"),
                   date_labels = "%b %d - %H:%M",
                   limits = c(as.POSIXct("0020-01-05 11:00:00"),
                              as.POSIXct("0020-01-8 16:00:00")))+
  xlab("")+
  ylab(expression(paste("Dew Point (",degree,"C )")))+
  ggtitle("Dew Point")+
   annotate("rect",xmin=as.POSIXct("0020-01-06 10:13:00"),
            xmax=as.POSIXct("0020-01-06 10:53:00"),
            ymin=-Inf,ymax=Inf, alpha=0.3, fill="#F8766D")+
   annotate("rect",xmin=as.POSIXct("0020-01-07 09:08:00"),
            xmax=as.POSIXct("0020-01-07 09:25:00"),
            ymin=-Inf,ymax=Inf, alpha=0.3, fill="#F8766D")+
   annotate("rect",xmin=as.POSIXct("0020-01-08 09:25:00"),
            xmax=as.POSIXct("0020-01-08 09:47:00"),
            ymin=-Inf,ymax=Inf, alpha=0.3, fill="#F8766D")+
   annotate("rect",xmin=as.POSIXct("0020-01-07 09:25:00"),
            xmax=as.POSIXct("0020-01-07 14:19"),
            ymin=-Inf,ymax=Inf, alpha=0.3, fill="#00BFC4")+
   annotate("rect",xmin=as.POSIXct("0020-01-08 09:47:00"),
            xmax=as.POSIXct("0020-01-08 15:11:00"),
            ymin=-Inf,ymax=Inf, alpha=0.3, fill="#00BFC4")+
theme(axis.text.x = element_text(size=12),axis.text.y = element_text(size=12),axis.title.y = element_text(size=12),title = element_text(size=14))+
  geom_point()

# arrange plots
ggarrange(lw,sr,tc,rh,dp, ncol=1,nrow=5,align = "v") -> FigureS1

#clean-up environment
rm(df_enviro,df_enviro2,dp,lw,rh,sr,tc)

```

# 3. Supply chain studies

## Plot bacteria counts by stage and lot
```{r, warning=FALSE}
df_micro_supplychain <- read_csv("Data/SupplyChainMicroData.csv")

df_micro_supplychain$Lot <- as.factor(df_micro_supplychain$Lot)
# prep data for plots (spell out stage and test levels in full & capitalize lot)
df_micro_supplychain$Stage <- factor(df_micro_supplychain$stage, 
                                     levels=c("H", "T", "P","D0","R","DI"), 
                                     labels=c("Harvest", 
                                              "Transit", 
                                              "Pre-packaging",
                                               "Packaging",
                                              "Local retail", 
                                              "Beijing delivery"))
df_micro_supplychain$Test <- factor(df_micro_supplychain$test, 
                                    levels=c("APC", "GN"), 
                                    labels=c("Aerobic plate count", 
                                             "Total Gram-negative count"))

# plot counts by lot and stage
ggplot(df_micro_supplychain, aes(x=Lot,y=log10count,col=Test))+
  geom_point(shape=1)+
  scale_y_continuous(limits=c(5,8),breaks = seq(5,8,1))+
  ylab(expression(paste("Bacteria levels (",log[10]," CFU/g)")))+
  xlab("Lot")+
  facet_grid(.~Stage)+ theme(legend.position = "none") -> FigureS3
```

## Kruskal-Wallis tests for rank by lot
```{r}
# prep data
df_micro_supplychain %>%
  filter(test=="APC") -> df_APC_sc
df_micro_supplychain %>%
    filter(test=="GN") -> df_GN_sc

# run Kruskal-Wallis for each stage and test
lapply(split(df_APC_sc, df_APC_sc$Stage), 
       function(d) { kruskal.test(log10count ~ Lot, data=d) })# all P >0.05

lapply(split(df_GN_sc, df_GN_sc$Stage), 
       function(d) { kruskal.test(log10count ~ Lot, data=d) }) # all P >0.05

# don't run post-hoc Dunn's test because K-W showed P>0.05 for all stages
```

## Summary stats
```{r}
# Summarize data for 9 samples from supply chain studies 
#(combining the 3 lots based on finding above)
df_micro_supplychain %>%
  group_by(test,Stage)%>%
  summarize(mean=mean(log10count),
            sd=sd(log10count),
            median=median(log10count),
            quantile_25=quantile(log10count,0.25),
            quantile_75=quantile(log10count,0.75),
            min=min(log10count),
            max=max(log10count),
            n=n()) -> TableS2
```

## Kruskal-Wallis tests for rank by stage (sep for APC & GN)
```{r}
# run Kruskal-Wallis test
kruskal.test(log10count ~ Stage, data=df_APC_sc) #P > 0.05

kruskal.test(log10count ~ Stage, data=df_GN_sc) #P > 0.05

# don't run post-hoc Dunn's test because K-W showed P>0.05 
```

## Plot bacteria counts by sample_rep for harvest & packaging
```{r}
## data prep
# make sure sample_rep is an unordered factor
df_micro_supplychain$sample_rep <- 
  factor(df_micro_supplychain$sample_rep, ordered=FALSE)

# harvest
df_APC_sc %>%
  filter(stage=="H")-> df_APC_sc_H
df_GN_sc %>%
  filter(stage=="H")-> df_GN_sc_H

# packaging
df_APC_sc %>%
  filter(stage=="D0")-> df_APC_sc_D0
df_GN_sc %>%
  filter(stage=="D0")-> df_GN_sc_D0

## plot
# harvest
h_apc <- ggplot(data=df_APC_sc_H, aes(x=sample_rep, y=log10count, col=Lot))+
  geom_point()+
  ylab(expression(paste("Aerobic plate count (",log[10]," CFU/g)")))+
  xlab("Replicate")+
  facet_grid(.~Stage)
h_gn <- ggplot(data=df_GN_sc_H, aes(x=sample_rep, y=log10count, col=Lot))+
  geom_point()+
  ylab(expression(paste("Total Gram-negative count (",log[10]," CFU/g)")))+
  xlab("Replicate")+
  facet_grid(.~Stage)

# packaging
d0_apc<- ggplot(data=df_APC_sc_D0, aes(x=sample_rep, y=log10count, col=Lot))+
  geom_point()+
  ylab(expression(paste("Aerobic plate count (",log[10]," CFU/g)")))+
  xlab("Replicate")+
  facet_grid(.~Stage)
d0_gn <- ggplot(data=df_GN_sc_D0, aes(x=sample_rep, y=log10count, col=Lot))+
  geom_point()+
  ylab(expression(paste("Total Gram-negative count (",log[10]," CFU/g)")))+
  xlab("Replicate")+
  facet_grid(.~Stage)

# arrange plots
ggarrange(h_apc,h_gn,d0_apc,d0_gn, ncol=2,nrow=2,common.legend = TRUE) -> FigureS4

# clean-up environment
rm(h_apc,h_gn,d0_apc,d0_gn)
```

## Kruskal-Wallis tests for rank by sample_rep
```{r}
## run Kruskal-Wallis rank by sample_rep
# harvest
kruskal.test(log10count ~ sample_rep, data=df_APC_sc_H) # marginally significant (P=0.066)
kruskal.test(log10count ~ sample_rep, data=df_GN_sc_H) # marginally significant (P=0.066)

# packaging
kruskal.test(log10count ~ sample_rep, data=df_APC_sc_D0) # P = 0.875 so don't perform post-hoc Dunn's test
kruskal.test(log10count ~ sample_rep, data=df_GN_sc_D0) # P = 0.957 so don't perform post-hoc Dunn's test
```

## Clean-up environment 
```{r}
# clean-up environment
rm(df_APC_sc, df_APC_sc_D0, df_APC_sc_H, df_GN_sc, df_GN_sc_D0, df_GN_sc_H)
rm(df_micro_supplychain)
```

# 4. Shelf life studies

## Data visualization by day, lot, location, and test
```{r}
#load data
df_micro_shelflife <- read_csv("Data/ShelfLifeMicroData.csv")

## Plot means over time for L shelf life samples
# prep data
df_micro_shelflife$Test <- df_micro_shelflife$test
df_micro_shelflife_L <- df_micro_shelflife %>%filter(location=="L")

# plot
df_micro_shelflife_L %>%
  filter(sl_day!=4|lot!="L2")%>% # exclude data from sl_day 4 for L2 for L for preparing means/lines for figure since 1 datapoint & apparent outlier
  group_by(sl_day,lot,Test)%>%
   summarize(mean_LOG10N=mean(log10count),median_LOG10N=median(log10count)) %>%
  ggplot(aes(x=sl_day,y=mean_LOG10N, group=Test))+
    geom_point(shape=19)+
    geom_point(data=df_micro_shelflife_L, aes(x=sl_day,y=log10count, col=Test), shape=1)+
    geom_line(aes(linetype=Test))+
    #geom_line()+
    #geom_line(aes(x=sl_day,y=median_LOG10N),linetype="dashed")+
    ylim(5,8)+scale_x_continuous(limits=c(0,10),breaks = seq(0,10,2))+scale_shape_manual(values=c(0,1))+
    xlab("Day postprocessing")+ ylab(expression(paste("Bacteria levels (",log[10]," CFU/g)")))+
    ggtitle("Local grocery")+theme(legend.position = "none")+facet_grid(.~lot) -> p_SL_L

## Plot means over time for E shelf life samples
# prep data
temp <- df_micro_shelflife %>% filter(sl_day==0)
temp$location <- "E"
df_micro_shelflife2 <- rbind(temp, df_micro_shelflife)
df_micro_shelflife_E <- df_micro_shelflife2 %>%  filter(location=="E")

# plot
df_micro_shelflife2 %>%
  filter(location=="E")%>%
  group_by(sl_day,lot,Test)%>%
   summarize(mean_LOG10N=mean(log10count),median_LOG10N=median(log10count)) %>%
  ggplot(aes(x=sl_day,y=mean_LOG10N, group=Test))+
    geom_point(shape=19)+
    geom_point(data=df_micro_shelflife_E, aes(x=sl_day,y=log10count, col=Test), shape=1)+
     geom_line(aes(linetype=Test))+
    #geom_line()+
    #geom_line(aes(x=sl_day,y=median_LOG10N),linetype="dashed")+
    ylim(5,8)+scale_x_continuous(limits=c(0,10),breaks = seq(0,10,2))+scale_shape_manual(values=c(0,1))+
    xlab("Day postprocessing")+ ylab(expression(paste("Bacteria levels (",log[10]," CFU/g)")))+
    ggtitle("eCommerce")+theme(legend.position = "none")+facet_grid(.~lot) -> p_SL_E

## arrange the 2 plots
ggarrange(p_SL_L,p_SL_E, nrow=2) -> Figure2

# clean-up environment
rm(temp,df_micro_shelflife2,p_SL_L,p_SL_E,df_micro_shelflife_E,df_micro_shelflife_L)
```

## Summary stats; Table 7
```{r}
# Summarize data for samples by lot from shelf life studies
TableS3 <- df_micro_shelflife %>%
  group_by(test,sl_day, location, lot) %>%
  summarize(mean=mean(log10count),
            min=min(log10count),
            max=max(log10count),
            n=n())
```

## Estimate growth rates for each location/test/lot & summarize
```{r}
#first, reviewed plots to determine:
# min mean is end of "lag phase"/start of "exponential phase" for L is 2 for all; "lag" for E is 4 for all
# max mean is end of "exponential phase"; day 6 for all L; day 8 for L1 and L3 E and day 6 for L2 E
#these ended up informing subset of "exponential phase"

# prep data for L (APC and GN); days 2 to 6
df_SL_L1_APC_L <- df_micro_shelflife %>% 
  filter(location=="L"&test=="APC"&lot=="L1")%>%
  filter(sl_day!=10&sl_day!=8&sl_day!=0)

df_SL_L2_APC_L <- df_micro_shelflife %>% 
  filter(location=="L"&test=="APC"&lot=="L2")%>%
  filter(sl_day!=10&sl_day!=8&sl_day!=0&sl_day!=4)# remove sl_day 4 outlier

df_SL_L3_APC_L <- df_micro_shelflife %>% 
  filter(location=="L"&test=="APC"&lot=="L3")%>%
  filter(sl_day!=10&sl_day!=8&sl_day!=0)

df_SL_L1_GN_L <- df_micro_shelflife %>% 
  filter(location=="L"&test=="GN"&lot=="L1")%>%
  filter(sl_day!=10&sl_day!=8&sl_day!=0)

df_SL_L2_GN_L <- df_micro_shelflife %>% 
  filter(location=="L"&test=="GN"&lot=="L2")%>%
  filter(sl_day!=10&sl_day!=8&sl_day!=0&sl_day!=4)# remove sl_day 4 outlier

df_SL_L3_GN_L <- df_micro_shelflife %>% 
  filter(location=="L"&test=="GN"&lot=="L3")%>%
  filter(sl_day!=10&sl_day!=8&sl_day!=0)

# prep data for E (APC and GN); days 4 to 8
df_SL_L1_APC_E <- df_micro_shelflife %>% 
  filter(location=="E"&test=="APC"&lot=="L1")%>%
  filter(sl_day!=10&sl_day!=3&sl_day!=0)

df_SL_L2_APC_E <- df_micro_shelflife %>%
  filter(location=="E"&test=="APC"&lot=="L2")%>%
  filter(sl_day!=10&sl_day!=3&sl_day!=0&sl_day!=8)# remove sl_day 8 outlier

df_SL_L3_APC_E <- df_micro_shelflife %>% 
  filter(location=="E"&test=="APC"&lot=="L3")%>%
  filter(sl_day!=10&sl_day!=3&sl_day!=0)

df_SL_L1_GN_E <- df_micro_shelflife %>% 
  filter(location=="E"&test=="GN"&lot=="L1")%>%
  filter(sl_day!=10&sl_day!=3&sl_day!=0)

df_SL_L2_GN_E <- df_micro_shelflife %>% 
  filter(location=="E"&test=="GN"&lot=="L2")%>%
  filter(sl_day!=10&sl_day!=3&sl_day!=0&sl_day!=8)# remove sl_day 8 outlier

df_SL_L3_GN_E <- df_micro_shelflife %>% 
  filter(location=="E"&test=="GN"&lot=="L3")%>%
  filter(sl_day!=10&sl_day!=3&sl_day!=0)

# determine estimated/approximate growth rate (in "exponential phase") for each unique lot/test/location 
rbind(df_SL_L1_APC_L[c(1:5,7:8)],df_SL_L2_APC_L[c(1:5,7:8)], df_SL_L3_APC_L[c(1:5,7:8)],
      df_SL_L1_GN_L[c(1:5,7:8)],df_SL_L2_GN_L[c(1:5,7:8)], df_SL_L3_GN_L[c(1:5,7:8)],
      df_SL_L1_APC_E[c(1:5,7:8)],df_SL_L2_APC_E[c(1:5,7:8)], df_SL_L3_APC_E[c(1:5,7:8)],
      df_SL_L1_GN_E[c(1:5,7:8)],df_SL_L2_GN_E[c(1:5,7:8)], df_SL_L3_GN_E[c(1:5,7:8)])-> temp

temp2 <- temp %>%
  group_by(test,sl_day, location, lot) %>%
  summarize(mean=mean(log10count))

gr.df <- temp2 %>%
  group_by(test,location, lot) %>%
  summarize(min=min(mean),
            max=max(mean),
            diff=max-min)
gr.df$d <- 4
gr.df$d[gr.df$location == "E"&gr.df$lot=="L2"]<-2
gr.df$gr <- gr.df$diff/gr.df$d

gr.df %>%
  group_by(test, location)%>%
  summarize(mean_growthrate=mean(gr),
            min_growthrate=min(gr),
            max_growthrate=max(gr))-> tbl_growthrates
# clean-up environment
rm(df_SL_L1_APC_L,df_SL_L2_APC_L,df_SL_L3_APC_L,df_SL_L1_GN_L,df_SL_L2_GN_L,df_SL_L3_GN_L,
   df_SL_L1_APC_E,df_SL_L2_APC_E,df_SL_L3_APC_E,df_SL_L1_GN_E,df_SL_L2_GN_E,df_SL_L3_GN_E)
rm(temp,temp2,df_micro_shelflife)
```

## Determine if growth rates differ by site and bacteria
```{r}
# fit linear regression model with data from above to determine whether growth rates differ by site and bacteria
gr.df$location <- factor(gr.df$location, levels = c("L","E"))

lm(gr ~ location, data=gr.df) -> gr.m1 # P < 0.05
summary(gr.m1)

lm(gr ~ test, data=gr.df) -> gr.m2 # P > 0.05
summary(gr.m2)

# final model is gr.m1 (Table3)
plot(gr.m1)
```


